Robust Adaptation to Multi-Scale Climate Variability

### 2019-03-29 report-earths-future/reviewer-response.md

# Response to Reviewers

We thank Dr. Montanari and one anonymous reviewer for their helpful critique of our paper "Robust Adaptation to Multi-Scale Climate Variability".
Major comments focused on the key points, the Hurst phenomenon, and connections between the introduction, results, and discussion.
We draw the attention of the reviewers to section 1.2, which has been completely re-written (the track changes package was not used because the section was completely re-organized) to include a more thorough theoretical discussion of the Hurst phenomenon, low-frequency variability, and our experimental design.

We also draw the attention of the reviewers to section 2.4, which now includes a diagram and some additional text to contextualize the experiment design.
THis section has also been completely re-written and we have also neglected to track changes in this section.

We also thank the reviewers for thoughtful specific comments.
We have incorporated nearly all of the changes as suggested; any discrepancies are discussed below.

@@

I would like to point out that I missed in the paper a discussion on the Hurst effect.
There is an extended literature on it. Such effect implies the presence of low frequency variability (long term cycles) which were observed in several secular time series of environmental variables.
The Hurst effect may be difficult to distinguish from anthropogenic climate change (ACC).
This effect is ignored by several climate studies. I believe this paper would be a good opportunity to discuss the interaction of the Hurst effect with ACC.
To this end, a proper discussion should be included in the manuscript.

I would like to congratulate the authors for their interesting work.

Best regards,
Alberto Montanari

> Thank you for your thoughtful and positive response. We share responses to your specific (line) comments below. We would like to point your attention to section 1.2, which has been substantially re-written and now provides a greater theoretical discussion of the Hurst phenomenon and low-frequency climate variability. Though this is a very deep literature and

@@

The authors have teased apart climatic risks that are considered for planning purposes across multiple time horizons.
It is an interesting exploration the interaction of various types of climatic variability, planning horizons, and length of time series, and how this interaction impacts the choice of instruments of climate risk mitigation.
As this paper uses the word robust in the title, the introduction needs to link the concept of robustness to their questions and methods.
Suggest citing literature on robust adaptation if you keep ‘robust’ in the title.

> See Line 39 (verify)

In Sections 2.1 and 2.2 you describe the three models of climatic change, and three ways of modeling each climate model.
Is there a way to clarify the structure of your computational experiment in these two sections, including your intent of using multiple methods of generating these types of climate variability?

> We have re-named section 2.4 experiment design and added a flow chart plus some clarifying text to this section.

@@

Please explicitly link the lognormal model with the LN2 acronym in text before using in figures.

> Thank you for noticing this omission. We have defined this acronym in section 2.2

@@

> We have added some modifications to the discussion to reference our computational results more directly and clearly. Please see underlined changes in section 4.

### Key points

To me, planning implies a process rather than the expected lifespan of a project.
Can you clarify which meaning you use?
If the latter, perhaps the following would be more readable: Climate risk varies over the planned lifespan of a project.
This can also be clarified in other locations of the paper, e.g. line 364 in the summary.

> Thank you for your comment on the key points. Based on suggestions from both reviewers we have completely re-written the key points.

### 2019-03-29 report-earths-future/paper-2018-robust-adaptation-cyclical-risk.tex
incorporate scott suggestions

% Other package calls
\usepackage{
  amssymb, amsmath,   % for math symbols
  apacite,            % AGU style citations apparently
  booktabs,           % for nice tables
  graphicx,           % for figures
  natbib,             % for references
  physics,            % for physics notation
  siunitx,            % for SI notation
  trackchanges,       % to track any new text
  xspace,             % for ie and eg
}

@@

% Use the glossaries package
\usepackage[acronym]{glossaries}
\makeglossaries
\newacronym{acc}{ACC}{anthropogenic climate change}
\newacronym{amo}{AMO}{Atlantic Multidecadal Oscillation}
\newacronym{cba}{CBA}{cost-benefit analysis}
\newacronym{enso}{ENSO}{the El Ni\~{n}o-Southern Oscillation}
\newacronym{ffa}{FFA}{flood frequency analysis}
\newacronym{gcm}{GCM}{general circulation model}
\newacronym{hmm}{HMM}{hidden Markov model}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
\newacronym{ipo}{IPO}{Interdecadal Pacific Oscillation}
\newacronym{lbda}{LBDA}{living blended drought analysis}
\newacronym{lfv}{LFV}{low-frequency variability}
\newacronym{nao}{NAO}{North Atlantic Oscillation}
\newacronym{pdo}{PDO}{Pacific Decadal Oscillation}
\newacronym{s2d}{S2D}{seasonal to decadal}
\newacronym{s2s}{S2S}{sub-seasonal to seasonal}

@@

Recent climate extremes such as floods, droughts, hurricanes, tornadoes, hailstorms, and heat waves have caused death and destruction, motivating investments in climate adaptation for the public and private sectors.
Further, rapid and continuing changes to global climate hazard and exposure underscore the need for adaptation strategies.
For example, population growth and urbanization have driven rapid increases in global exposure to events such as floods \citep{Jongman:2012cr} and tropical cyclones \citep{Peduzzi:2012iq}.
At the same time, anthropogenic modification of global and local climate processes affects the frequency, intensity, and location of extreme events \citep{IPCC:2012wt, Milly:2008dg, Shaw:2016bo}.
Even if \add{future} mitigation efforts are successful, \add{existing levels of atmospheric $\text{CO}_2$ and ocean heat content mean that} novel adaptation strategies are needed.

This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, e.g.\@\xspace,][]{Merz:2014gf}.
\add{A typical goal is to create systems which are robust in the sense that they perform well over a wide range of plausible futures} \citep{lempert:2007,Borgomeo:2018hk} \add{and which fail along non-catastrophic modes} \citep{brown:2010}.
Although climate risk has been traditionally managed with centrally planned structural instruments (e.g.\@\xspace, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently slowed their implementation.
\remove{Instead}\add{Rather}, factors such as New York City have turned to a combination of structural\add{ (e.g.\@\xspace, a stormwater barrier)}, operational\add{ (e.g.\@\xspace, improved evacuation routes)}, and financial\add{ (e.g.\@\xspace, index insurance)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
These instruments are not typically implemented in isolation or statically.
Instead, investment decisions made at each point in time affect the viability, costs, and benefits of future decisions, causing the system to trace a ``pathway'' through time \citep{Walker:2013gi,Haasnoot:2013im,Haasnoot:2015dh}.

Despite recent insights, important questions remain.
How should a portfolio of risk mitigation instruments be optimized?
How should one choose between permanent and transient instruments?
Under what conditions is a permanent, large infrastructure investment required, and what information is needed to recognize this threshold?
In this paper we focus on \add{the more narrow question of} how the temporal structure of climate risk, and the uncertainties associated with \add{its estimation}\remove{estimating it}, influence the answers to these questions.
We continue this section with three specific observations about climate risk which, while apparently obvious, have important and subtle implications\add{,} which we examine in \cref{sec:methods,sec:results,sec:discussion-conclusions}.

@@

Climate risk is governed by a variety of physical processes which occur on scales ranging from local and transient to global and permanent.
Of these processes, \gls{acc} has received the most attention in the climate adaptation literature and its influence on \add{some} river floods, droughts, hurricanes, urban flooding, and many other climate hazards \remove{are}\add{has been the subject of substantial investigation} \remove{well documented} \citep[e.g.\@\xspace,][]{Coumou:2012bc,Milly:2008dg,OGorman:2009hj,Trenberth:2003bj}.
Human activities can also affect climate risk through modification of local land or river systems \citep[see][]{Merz:2014gf}, and through changes in exposure to extremes \citep{baldassarre:2018,Jongman:2012cr}.
In combination, these effects highlight that the past may not be an adequate representation of future climate risk \citep[termed ``nonstationarity'' by][]{Milly:2008dg}.

Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.
The Hurst phenomenon associated with stochastic long-memory processes has been invoked to explain long scale fluctuations in geophysics, physics, biology, medicine, traffic, network dynamics as well as for financial time series:
\begin{equation}
  \frac{R_n}{S_n} \sim n^h,
\end{equation}
where $h \in [0.5, 1]$, $R_n$ is the range of the first $n$ data points, and $S_n$ is the sample standard deviation \citep{Hurst:1951tc,oconnell:2016}.
A variety of theories, including infinite memory, nonstationarity of the process, and specific storage mechanisms have been put forth to explain the Hurst phenomenon, none fully satisfactorily \citep{klemes:1974}.
The Hurst exponent is directly related to the fractal dimension of a process, and there is a rich multi-disciplinary literature as to the process level and statistical justification of long memory and fractal processes, especially in hydrology \citep{mandelbrot:1985,mandelbrot:1969,beran:1994}.
\citet{lovejoy:2012,lovejoy:2013a} and \citet{lovejoy:2013} also argue that multifractal and long memory processes are an appropriate description of climate in general and weather extremes in particular. 
Based on this theory, regimes and quasi-periodic phenomena with increasing amplitude as their recurrence interval increases would be the norm rather than the exception.
Similar arguments are made by \citet{koutsoyiannis:2003}, \citet{markonis:2013}, and \citet{oconnell:2016}, and detailed discussions and some philosophical perspectives are offered by \citet{bloschl:2010} and \citet{montanari:2003}.
Wavelet methods have also been used to estimate the Hurst exponent \citep{simonsen:1998,chamoli:2007}, and to design simulation algorithms that reproduce self-similarity, long range dependence and quasi-periodic regimes \citep{Kwon:2007fj,bullmore:2001,geweke:1983,feng:2005}.

Links between local hydrological processes and large-scale climate patterns, often referred to as teleconnections \citep{Angstrom:1935ej}, have also been used to explain \gls{lfv} in hydroclimate systems.
If these climate modes vary on decadal or longer time scales, such as the Pacific Decadal Oscillation \citep{Mantua:1997kj}, then this \gls{lfv} may appear in local hydrologgical records as well, leading the past to be a poor predictor of future behavior.
For example, \citet{Jain:2001hz} demonstrated that in a system with substantial \gls{lfv}, due, e.g.\@\xspace, to \gls{enso}, the past $N$ years can be a poor proxy for the following $N$ years even when no secular change is present.
When records are short, and whether low-frequency external forcing or local memory dominates, it is easy to mistake the Hurst phenomenon for a secular trend which may not continue from the past into the future \citep{bhattacharya:1983}.

We illustrate this type of \gls{lfv} in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda} \citep{Cook:2010bz}, a \SI{100}{year} record of daily river stage on the American River at Folsom, and the wavelet average spectrum for both \citep{Torrence:1998jp,Roesch:wlBQQoIs}.
Peaks for the American River time series are apparent at approximately 2 and 15 years and in the \gls{lbda} time series at approximately 8, 20, and 64 years.
This is illustrated by the blue line in the top left panel of \cref{fig:observed-lfv}, which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
A detailed analysis of these time series is beyond the scope of this paper, but we note that the high amplitude and long time periods of the quasi-periodic oscillations they exhibit are consistent with analyses of \gls{lfv} in other hydroclimate systems \citep{Kiem:2002kq,Swierczynski:2012km,Woollings:2014kd,Hodgkins:2017hw}.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{observed-lfv.pdf}
  \caption{
    Hydroclimate time series vary on many time scales.
    (a) A \SI{500}{year} reconstruction of summer rainfall over Arizona from the \acrfull{lbda}.
    High numbers indicate more severe drought.
    A 20-year running mean is also shown in blue.
    (b) A \SI{100}{year} record of annual-maximum streamflows for the American River at Folsom.
    (c) the wavelet global (average) spectrum of the \gls{lbda} time series (a).
    Blue dots indicate frequencies which are significant at $\alpha=0.10$ and red dots frequencies which are significant at $\alpha=0.05$ compared to white noise.
    (d) wavelet spectrum, like (c), for the American River data.
  }\label{fig:observed-lfv}
\end{figure}

@@

Evaluating a particular risk mitigation instrument involves projecting climate risk over the $M$-year planning period.
Consequently, the physical mechanisms which impart predictability on the system differ between projects with long and short planning periods.
As illustrated in \cref{fig:conceptual-sketch} (a), the lifetime risk of a permanent structure with a \SI{100}{year} planning period depends on the magnitude and extent of future human activities, with very large associated uncertainty.
Even \remove{with} \add{in the idealized, and unrealistic, case of} a perfect climate model, these uncertainties will be large.
By contrast, this perfect climate model may usefully inform estimates of climate hazard over a three-year insurance contract with much less associated uncertainty.

@@

The first step is to sample climate risk by generating synthetic streamflow sequences.
To do this, we model  annual-maximum flood peaks with a log-normal distribution, conditional on a location parameter which varies in time:
\begin{equation} \label{eq:log-normal}
  \log Q(t) \sim \mathcal{N} \qty(\mu(t), \sigma(t)).
\end{equation}
We further assume a constant coefficient of variation of the log streamflow,
\begin{equation}
  \sigma(t) = \xi \mu(t)  
\end{equation}
and apply a lower threshold on the standard deviation
\begin{equation}
  \sigma(t) \geq \sigma_\text{min} > 0.
\end{equation}
This formulation describes the \begin{enumerate*}[label= (\roman*)]
  \item \gls{lfv} only;
  \item \gls{lfv} plus secular change; and
  \item secular change only
\end{enumerate*}
scenarios in a single equation; the only component which needs to change is the dependence of $\mu(t)$ on time, which we parameterize as
\begin{equation}\label{eq:nino3}
  \mu(t) = \mu_0 + \beta x(t) + \gamma \qty(t - t_0),
\end{equation}
where $x(t)$ represents a climate time series which itself exhibits \gls{lfv} but not secular change.
This parameterization is analagous to the ``climate-informed'' approach described in several studies for estimating climate risk \citep{Delgado:2014ey,Merz:2014gf,Farnham:2018gs}.
Following \cref{eq:nino3}, when $\beta \neq 0$ there will be \gls{lfv}, and when $\gamma \neq 0$ there will be secular change.
The values of all parameters used for sampling climate risk are listed in \cref{sec:supp-methods} for each of the three scenarios considered.

We represent the climate state variable $x(t)$ through an index for \gls{enso}, which has been shown to impact flood risk around the world \citep{Ropelewski:1987do, Ward:2014gg} and has characteristic variability on timescales of 3 to 7 years \citep{Sarachik:2009dr} as well as a ``staircase'' of lower-frequency scales \citep{Jin:1994wq}.
We model \gls{enso} variability by taking a \SI{20000}{year} integration of the Cane-Zebiak model \citep{Zebiak:1987cl} to produce a monthly NINO3 index \citep{Ramesh:2016hf}.
To create an annual time series, we average the October-December values of the NINO3 index for each year.
\Cref{sec:supp-nino-spectrum} shows a wavelet spectrum and time series plot of the resulting annual time series.
In \cref{sec:markov-generating} we consider an alternative parameterization of $\mu(t)$, i.e.\@\xspace an alternative to \cref{eq:nino3}, which considers a Markovian state transition rather than an explicit \gls{enso} model, and note a general agreement of results.

@@

Once a synthetic streamflow sequence has been generated, we evaluate the identifiability and predictability of the dominant climate modes by fitting the sequence to statistical models and creating probabilistic projections of the future.
We use three well-studied statistical methods for future flood risk, each of which parameterizes time in a different way.
One is purely stationary, another captures \gls{lfv}, and the third captures secular change.
We choose these models for their interpretability and simplicity, rather than because of a belief that they are generally valid.
For each synthetic flood sequence to be analyzed, the first $N$ years are treated as observations.
Once a statistical model is fit to these ``observations'', then $K=\num{1000}$ sequences of future annual-maximum streamflow over the future $M$-year record are generated from the fitted model using Monte Carlo simulation.

\add{In the first case we} \remove{We first consider fitting}\add{fit} a stationary model to the observed flood record, following classical assumptions of \gls{iid} sequences.
In this model annual-maximum streamflow are taken to follow a log-normal distribution with constant mean and variance.
\add{We refer to this model as ``LN2 Stationary.''}
The parameters of the model are fit  in a Bayesian framework to fully represent the posterior uncertainty, using the stan probabilistic computing package \citep{Carpenter:2017ke} with weakly informative priors \citep{gelman:2017,simpson:2017}.
The full model, including priors, is given in \cref{eq:log-normal-stationary}.

Next, we modify this stationary model to incorporate secular change.
Many studies have done this by regressing certain parameters of the model on time \citep[see][for a comprehensive review]{Salas:2018ge}.
We consider an extension of the stationary log-normal model by adding a time trend on the scale parameter and maintaining a constant coefficient of variation, as given in \cref{eq:log-normal-trend}.
\add{We refer tho this model as ``LN2 Linear Trend''}
This model gives a lower bound on total informational uncertainty because it correctly represents the trend's known form, whereas in real-world analyses the form of the trend is unknown.

\remove{We finally consider}\add{Finally, we explicitly model} \remove{a HMM which explicitly represents}\gls{lfv} \add{using a} \gls{hmm}.
A \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (i.e.\@\xspace hidden) states $S(t)$ \citep{Rabiner:1986jk}.
The (unobserved) states evolve following a first-order Markov process, and the observed variable (e.g.\@\xspace streamflow) depends only on the underlying state.
\Glspl{hmm} have been widely used for modeling streamflow sequences \citep{Bracken:2016ba} and \gls{enso} \citep{rojohernandez:2017}.
We fit  streamflow sequences using a \gls{hmm} with two states.
The model is fit using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution \remove{conditional only on the state}\add{that is conditional only on the unobserved state variables}.
This algorithm simultaneously estimates the transition matrix of the Markov process and the conditional parameters of each distribution.
For simplicity, we fit only a two-state \gls{hmm} to each sequence.
Future floods are then estimated by simulating future states from the estimated transition matrix and then drawing $Q(t)$ conditional on the simulated state.

@@

Both estimation bias and estimation uncertainty affect the utility of a climate risk projection, as illustrated in \cref{fig:conceptual-bias-variance}.
An instrument \add{whose design was based on projections with} \remove{designed based on} overestimated variance \remove{and}\add{or} positive bias will be over-designed, either causing the risk manager to avoid the investment, given its higher cost, or will lead to unnecessary diversion of funds from other instruments.
Similarly, an instrument designed based on underestimated variance or negative bias may be under-designed, and thus fail to protect the user.
\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{Bias-Variance-Sketch.pdf}
  \caption{
    Consequences of model bias or incorrect model representation of uncertainty.
    If an estimate has a positive bias and overestimates uncertainty, the instrument may be too expensive for the user.
    If an estimate has negative bias and underestimates uncertainty, it will be likely to fail.
  }\label{fig:conceptual-bias-variance}
\end{figure}

>>>>

% Other package calls
\usepackage{
  amssymb, amsmath,   % for math symbols
  apacite,            % AGU style citations apparently
  booktabs,           % for nice tables
  graphicx,           % for figures
  natbib,             % for references
  physics,            % for physics notation
  siunitx,            % for SI notation
  xspace,             % for ie and eg
}

@@

% Use the glossaries package
\usepackage[acronym]{glossaries}
\makeglossaries
\newacronym{acc}{ACC}{anthropogenic climate change}
\newacronym{amo}{AMO}{Atlantic Multidecadal Oscillation}
\newacronym{cba}{CBA}{cost-benefit analysis}
\newacronym{enso}{ENSO}{the El Ni\~{n}o-Southern Oscillation}
\newacronym{ffa}{FFA}{flood frequency analysis}
\newacronym{gcm}{GCM}{general circulation model}
\newacronym{hmm}{HMM}{hidden Markov model}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
\newacronym{ipo}{IPO}{Interdecadal Pacific Oscillation}
\newacronym{lbda}{LBDA}{Living Blended Drought Analysis}
\newacronym{lfv}{LFV}{low-frequency variability}
\newacronym{nao}{NAO}{North Atlantic Oscillation}
\newacronym{pdo}{PDO}{Pacific Decadal Oscillation}
\newacronym{s2d}{S2D}{seasonal to decadal}
\newacronym{s2s}{S2S}{sub-seasonal to seasonal}

@@

Recent climate extremes such as floods, droughts, hurricanes, tornadoes, hailstorms, and heat waves have caused death and destruction, motivating investments in climate adaptation for the public and private sectors.
Further, rapid and continuing changes to global climate hazard and exposure underscore the need for adaptation strategies.
For example, population growth and urbanization have driven rapid increases in global exposure to events such as floods \citep{Jongman:2012cr} and tropical cyclones \citep{Peduzzi:2012iq}.
At the same time, anthropogenic modification of global and local climate processes affects the frequency, intensity, and location of extreme events \citep{IPCC:2012wt, Milly:2008dg, Shaw:2016bo}.
Even if \add{future} mitigation efforts are successful, \add{existing levels of atmospheric $\text{CO}_2$ and ocean heat content motivate the development of} novel adaptation strategies\remove{ are needed}.

This need has motivated a multitude of approaches for estimating the probability distribution of future climate risk, and for choosing between different risk mitigation instruments based on these estimates \citep[see, e.g.\@\xspace,][]{Merz:2014gf}.
\add{A typical goal is to create systems which are robust in the sense that they perform well over a wide range of plausible futures} \citep{lempert:2007,Borgomeo:2018hk} \add{and which fail along non-catastrophic modes} \citep{brown:2010}.
Although climate risk has been traditionally managed with centrally planned structural instruments (e.g.\@\xspace, a levee), the high price \citep{Papakonstantinou:2016ve}, environmental costs \citep{dugan:2010}, and vulnerability to biased climate projections \citep{lempert:2007} have recently slowed their implementation.
\remove{Instead}\add{Rather}, actors such as New York City have turned to a combination of structural\add{ (e.g.\@\xspace, stormwater barrier)}, operational\add{ (e.g.\@\xspace, improved evacuation routes)}, and financial\add{ (e.g.\@\xspace, index insurance)} instruments for reducing vulnerability and increasing resilience to climate extremes \citep{CityofNewYork:2013uh}.
These instruments are not typically implemented in isolation or statically.
Instead, investment decisions made at each point in time affect the viability, costs, and benefits of future decisions, causing the system to trace a ``pathway'' through time \citep{Walker:2013gi,Haasnoot:2013im,Haasnoot:2015dh}.

Despite recent insights, important questions remain.
How should a portfolio of risk mitigation instruments be optimized?
How should one choose between permanent and transient instruments?
Under what conditions is a permanent, large infrastructure investment required, and what information is needed to recognize this threshold?
In this paper we focus on \add{the narrower question of} how the temporal structure of climate risk, and the uncertainties associated with \add{its estimation}\remove{estimating it}, influence the answers to these questions.
We continue this section with three specific observations about climate risk which, while \remove{apparently}\add{seemingly} obvious, have important and subtle implications \add{that}\remove{which} we examine in \cref{sec:methods,sec:results,sec:discussion-conclusions}.

@@

Climate risk is governed by a variety of physical processes which occur on scales ranging from local and transient to global and permanent.
Of these processes, \gls{acc} has received the most attention in the climate adaptation literature and its influence on some river floods, droughts, hurricanes, urban flooding, and many other climate hazards has been the subject of substantial investigation \citep[e.g.\@\xspace,][]{Coumou:2012bc,Milly:2008dg,OGorman:2009hj,Trenberth:2003bj}.
Human activities can also affect climate risk through modification of local land or river systems \citep[see][]{Merz:2014gf}, and through changes in exposure to extremes \citep{baldassarre:2018,Jongman:2012cr}.
In combination, these effects highlight that the past may not be an adequate representation of future climate risk \citep[termed ``nonstationarity'' by][]{Milly:2008dg}.

Secular change is not the only mechanism which can cause historical records to provide a biased view of future risk.
The Hurst phenomenon associated with stochastic long-memory processes and has been invoked to explain long scale fluctuations in geophysics, physics, biology, medicine, traffic, network dynamics as well as for financial time series.
This phenomena can be characterized as:
\begin{equation}\label{eq:hurst}
  \frac{R_n}{S_n} \sim n^H,
\end{equation}
where $H \in [0.5, 1]$ is the Hurst coefficient, $R_n$ is the range of the first $n$ cumulative deviations from the mean, and $S_n$ is the sample standard deviation \citep{Hurst:1951tc,oconnell:2016,klemes:1974}.
\Cref{eq:hurst} measures the tendency of a time series to either regress to the mean or cluster towards the extremes.
The Hurst exponent $H$ is directly related to the fractal dimension of a process, and there is a rich multi-disciplinary literature as to the process level and statistical justification of long memory and fractal processes, especially in hydrology \citep{mandelbrot:1985,mandelbrot:1969,beran:1994}.
\citet{lovejoy:2012,lovejoy:2013a} and \citet{lovejoy:2013} also argue that multifractal and long memory processes are an appropriate description of climate in general and weather extremes in particular. 
Based on this theory, regimes and quasi-periodic phenomena with amplitude that increases with recurrence interval would be the norm rather than the exception.
Similar arguments are made by \citet{koutsoyiannis:2003}, \citet{markonis:2013}, and \citet{oconnell:2016}, and detailed discussions and some philosophical perspectives are offered by \citet{bloschl:2010} and \citet{montanari:2003}.
Wavelet methods have also been used to estimate the Hurst exponent \citep{simonsen:1998,chamoli:2007}, and to design simulation algorithms that reproduce self-similarity, long range dependence and quasi-periodic regimes \citep{Kwon:2007fj,bullmore:2001,geweke:1983,feng:2005}.

Links between local hydrological processes and large-scale climate patterns, often referred to as teleconnections \citep{Angstrom:1935ej}, have also been used to explain \gls{lfv} in hydroclimate systems.
If these climate modes vary on decadal or longer time scales, such as the \gls{pdo} \citep{Mantua:1997kj}, then this \gls{lfv} may appear in local hydrologgical records as well, leading the past to be a poor predictor of future behavior.
For example, \citet{Jain:2001hz} demonstrated that in a system with substantial \gls{lfv}, due to \gls{enso}, the past $N$ years can be a poor proxy for the following $N$ years even when no secular change is present.
When records are short and low-frequency external forcing or local memory dominates, it is easy to mistake the Hurst phenomenon for a secular trend which may not continue from the past into the future \citep{bhattacharya:1983}.
t is also difficult to identify the dominant scales of variability \citep{newman:2016,Jin:1994wq}.

We illustrate this type of \gls{lfv} in \cref{fig:observed-lfv}, which shows a \SI{500}{year} drought reconstruction from the \gls{lbda} \citep{Cook:2010bz}, a \SI{100}{year} record of annual maximum streamflowflow on the American River at Folsom, and the global wavelet power spectrum for both \citep{Torrence:1998jp,Roesch:wlBQQoIs}.
Peaks for the American River time series are apparent at approximately 2 and 15 years and in the \gls{lbda} time series at approximately 8, 20, and 64 years.
This is illustrated by the blue line in \remove{the top left panel of} \cref{fig:observed-lfv}\add{(b)}, which shows a \SI{20}{year} moving average of the \gls{lbda} time series.
A detailed analysis of these time series is beyond the scope of this paper, but we note that the high amplitude and long time periods of the quasi-periodic oscillations they exhibit are consistent with analyses of \gls{lfv} in other hydroclimate systems \citep{Kiem:2002kq,Swierczynski:2012km,Woollings:2014kd,Hodgkins:2017hw}.

\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{observed-lfv.pdf}
  \caption{
    Hydroclimate time series vary on many time scales.
    (a) A \SI{500}{year} reconstruction of summer rainfall over Arizona from the \acrfull{lbda}.
    \remove{High numbers}\add{Lower values} indicate more severe drought.
    A 20-year running mean is also shown in blue.
    (b) A \SI{100}{year} record of annual-maximum streamflow for the American River at Folsom.
    Daily streamflow values were divided by the catchment area to yield a normalized flow in units of \si{\milli\meter\per\day}.
    (c) The global wavelet power spectrum of the \gls{lbda} time series (a).
    Blue (red) dots indicate frequencies which are significant at $\alpha=0.10(0.05)$ compared to white noise.
    (d) Global wavelet power spectrum, like (c), for the American River data.
  }\label{fig:observed-lfv}
\end{figure}

@@

Evaluating a particular risk mitigation instrument involves projecting climate risk over the $M$-year planning period.
Consequently, the physical mechanisms which impart predictability on the system differ between projects with long and short planning periods.
As illustrated in \cref{fig:conceptual-sketch} (a), the lifetime risk of a permanent structure with a \SI{100}{year} planning period depends on the magnitude and extent of future human activities, with very large associated uncertainty.
Even \remove{with} \add{in the idealized and unrealistic case of} a perfect climate model, these uncertainties will be large.
By contrast, this perfect climate model may usefully inform estimates of climate hazard over a three-year insurance contract with much less associated uncertainty.

@@

The first step is to sample climate risk by generating synthetic streamflow sequences.
To do this, we model  annual-maximum flood peaks with a log-normal distribution, conditional on a location parameter which varies in time:
\begin{equation} \label{eq:log-normal}
  \log Q(t) \sim \mathcal{N} \qty(\mu(t), \sigma(t)).
\end{equation}
We further assume a constant coefficient of variation of the log streamflow,
\begin{equation}
  \sigma(t) = \xi \mu(t)  
\end{equation}
and apply a lower threshold on the standard deviation
\begin{equation}
  \sigma(t) \geq \sigma_\text{min} > 0.
\end{equation}
This formulation describes \add{all} scenarios \add{for future climate considered in this paper within} \remove{in} a single equation\remove{;}\add{.}
\add{To add climate variability to the system,} the only component which needs to change is the dependence of $\mu(t)$ on time, which we parameterize as
\begin{equation}\label{eq:nino3}
  \mu(t) = \mu_0 + \beta x(t) + \gamma \qty(t - t_0),
\end{equation}
where $x(t)$ represents a climate time series which itself exhibits \gls{lfv} but not secular change.
This parameterization is analagous to the ``climate-informed'' approach described in several studies for estimating climate risk \citep{Delgado:2014ey,Merz:2014gf,Farnham:2018gs}.
Following \cref{eq:nino3}, when $\beta \neq 0$ there will be \gls{lfv}, and when $\gamma \neq 0$ there will be secular change.
The values of all parameters used for sampling climate risk are listed in the online supporting information for each of the three scenarios considered.

We represent the climate state variable $x(t)$ through an index for \gls{enso}, which has been shown to impact flood risk around the world \citep{Ropelewski:1987do, Ward:2014gg} and has characteristic variability on timescales of 3 to 7 years \citep{Sarachik:2009dr} as well as a ``staircase'' of lower-frequency scales \citep{Jin:1994wq}.
We model \gls{enso} variability by taking a \SI{20000}{year} integration of the Cane-Zebiak model \citep{Zebiak:1987cl} to produce a monthly NINO3 index \citep{Ramesh:2016hf}.
To create an annual time series, we average the October-December values of the NINO3 index for each year.
Supplemental figure S1 shows a wavelet spectrum and time series plot of the resulting annual time series.
In the online supplemental information we consider an alternative parameterization of $\mu(t)$, \remove{i.e.\@\xspace an alternative to,} which considers a Markovian state transition rather than an explicit \gls{enso} model, and note a general agreement of results.

@@

Once a synthetic streamflow sequence has been generated, we evaluate the identifiability and predictability of the dominant climate modes by fitting the sequence to statistical models and creating probabilistic projections of the future.
We use three well-studied statistical methods for future flood risk, each of which parameterizes time in a different way.
One is purely stationary, another captures \gls{lfv}, and the third captures secular change.
We choose these models for their interpretability and simplicity, rather than because of a belief that they are generally valid.
For each synthetic flood sequence to be analyzed, the first $N$ years are treated as observations.
Once a statistical model is fit to these \remove{``}observations\remove{''}, then $K=\num{1000}$ sequences of future annual-maximum streamflow over the future $M$-year record are generated from the fitted model using Monte Carlo simulation.

\add{In the first case we} \remove{We first consider fitting}\add{fit} a stationary model to the observed flood record, following classical assumptions of \gls{iid} sequences.
In this model annual-maximum streamflow are taken to follow a log-normal distribution with constant mean and variance.
\add{We refer to this model as ``LN2 Stationary.''}
The parameters of the model are fit  in a Bayesian framework to fully represent the posterior uncertainty, using the stan probabilistic computing package \citep{Carpenter:2017ke} with weakly informative priors \citep{gelman:2017,simpson:2017}.
The full model, including priors, is given in supplemental equation (S1).

Next, we modify this stationary model to incorporate secular change.
Many studies have done this by regressing certain parameters of the model on time \citep[see][for a comprehensive review]{Salas:2018ge}.
We consider an extension of the stationary log-normal model by adding a time trend on the scale parameter and maintaining a constant coefficient of variation, as given in supplemental equation (S2).
\add{We refer to this model as ``LN2 Linear Trend''}
This model gives a lower bound on total informational uncertainty because it correctly represents the trend's known form, whereas in real-world analyses the form of the trend is unknown.

\remove{We finally consider}\add{Finally, we explicitly model} \remove{a HMM which explicitly represents}\gls{lfv} \add{using a} \gls{hmm}.
An \gls{hmm} is a latent variable model in which the system being modeled is assumed to follow a Markov process with unobserved (i.e.\@\xspace hidden) states $S(t)$ \citep{Rabiner:1986jk}.
The (unobserved) states evolve following a first-order Markov process, and the observed variable (e.g.\@\xspace streamflow) depends only on the underlying state.
\Glspl{hmm} have been widely used for modeling streamflow sequences \citep{Bracken:2016ba} and \gls{enso} \citep{rojohernandez:2017}.
We fit  streamflow sequences using a \gls{hmm} with two states.
The model is fit using the Baum-Welch algorithm, assuming that the data follow a log-normal distribution \remove{conditional only on the state}\add{that is conditional only on the unobserved state variables}.
This algorithm simultaneously estimates the transition matrix of the Markov process and the conditional parameters of each distribution.
For simplicity, we fit only a two-state \gls{hmm} to each sequence.
Future floods are then estimated by simulating future states from the estimated transition matrix and then drawing $Q(t)$ conditional on the simulated state.

@@

Both estimation bias and estimation uncertainty affect the utility of a climate risk projection, as illustrated in \cref{fig:conceptual-bias-variance}.
An instrument \add{whose design was based on projections with} \remove{designed based on} overestimated variance \remove{and}\add{or} positive bias will be over-designed, either causing the risk manager to avoid the investment, given its higher cost, or will lead to unnecessary diversion of funds from other instruments.
Similarly, an instrument designed based on underestimated variance or negative bias may be under-designed, and thus fail to protect the \remove{user}\add{public}.
\begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{Bias-Variance-Sketch.pdf}
  \caption{
    Consequences of model bias or incorrect model representation of uncertainty.
    If an estimate has a positive bias and overestimates uncertainty, the instrument may be too expensive.
    If an estimate has negative bias and underestimates uncertainty, it will be likely to fail.
  }\label{fig:conceptual-bias-variance}
\end{figure}
